%% -*- LaTeX -*-
\documentclass{report}

\usepackage{thesis}

%some other packages I found useful, comment out or remove if you do not need them
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[titletoc,toc,title]{appendix}
\usepackage{chapterbib}
\usepackage{cite}
\usepackage[pdftex]{graphicx}
\usepackage{mdframed}
\usepackage{minted}
\usepackage{multirow}
\usepackage{pdfcomment}
\usepackage{subfigure}
\usepackage{tabularx}
\usepackage{titlesec}
\usepackage{url}
\usepackage{xspace}

% must be loaded after hyperref
\usepackage{cleveref}

\newcommand{\inlinecode}[2][haskell]{\mintinline[breaklines]{#1}{#2}}

\newcommand{\ChoreographyTS}{ChoreoTS\xspace}
\newcommand{\chorLambda}{Chor$\lambda$\xspace}
\newcommand{\Chorus}{ChoRus\xspace}
\newcommand{\HasChor}{Has\-Chor\xspace}
\newcommand{\HLSCentral}{$\boldsymbol{\lambda}_{\boldsymbol{C}}$\xspace}
\newcommand{\HLSLocal}{$\boldsymbol{\lambda}_{\boldsymbol{L}}$\xspace}
\newcommand{\HLSNet}{$\boldsymbol{\lambda}_{\boldsymbol{N}}$\xspace}
\newcommand{\MultiChor}{Multi\-Chor\xspace}

\widowpenalty=1000000
\clubpenalty=1000000


\title{Choreographic Programming 2: Electric Boogalo}
\author{Mako Bates}
\defensedate{June XX, 2025}
\dissertation                             %% or \dissertation
\doctorphilosophy                      %% or \doctorphilosophy or \masterarts
\cs                                 %% this is the only speciality defined.
\graduatingIn{August}
\advisor{Bunny Lebowski, Ph.D.}
\readerone{Walter Sobchak, Ph.D.}
\readertwo{Carl Hungus, Ph.D.}      %% Optional if MS.
\chair{V.I. Lenin, Ph.D.}
\dean{Holger Hoock, Ph.D.}  %% TODO: check title.


\begin{document}
\maketitle
\makeacceptance

\begin{abstract}
Presents both MultiChor (and possibly a distinct variant MiniChor) and a formalism "He Lambda something".
Shows how this system is more ergonomic than prior work and equally as expressive.
Expolores the theoretical minimum of choreographic programming, and implications for real use.
\end{abstract}

\begin{citationspage}
\citationheadingpublished
\noindent Doe, J. and B. Lebowski. (2009). My First Published Paper. {\it Proceedings of the IEEE Congress on Life}.
\\ \\
\noindent Doe, J. and B. Lebowski. (2010). My Second Published Paper. {\it The World's Greatest Journal}.

%if have unpublished work, use one or more of the following
\begin{center}
\vskip 2em
{\large AND}
\end{center}
\vskip 2em 
Material from this dissertation has been accepted for publication in {\it Fantastic Journal} on 05/20/2013 in the following form:\\
\vskip 2em 
\noindent Doe, J. and B. Lebowski. (2013). Soon To Be Published Paper. {\it Fantastic Journal}.

\begin{center}
\vskip 2em
{\large AND}
\end{center}
\vskip 2em 
Material from this dissertation has been submitted for publication in {\it Some Other Journal} on 04/30/2013 in the following form:\\
\vskip 2em 
\noindent Doe, J. and B. Lebowski. (2013). Paper In Review. {\it Some Other Journal}.

\end{citationspage}

\begin{dedication} %% Optional
{\large 
{\it in memory of} 
\\
\vskip 2em
Alan Turing (1912-1954)}
\end{dedication}

\begin{acknowledgements}  %% Optional
I'd like to take this opportunity to pour a little of my 40oz. out for all
the homies that didn't make it.
\end{acknowledgements}

\tableofcontents
\listoffigures
\listoftables

\mainmatter
\sloppy

%for journal format dissertation each chapter needs its own reference section, and want references to be included in global bibliography
%this is accomplished using include commands
%but, for things like the lit review (chapter 1) or an appendix, should use input so it does not get it's own references section

%if not doing a journal format dissertation you can just use input for all chapters, or put text here directly.
\input{chapter1}
\include{chapter2}
\include{chapter3}

\bibliographystyle{chicago}
\bibliography{example}

\input{appendix}

\end{document}
